---
title: "Assignment 7"
author:
  name: Natalia
  date: 12.10.2025
format: html
execute:
  echo: true
  warning: false
  message: false
---

# Assignment 7

Use tidyr and dplyr libraries whenever possible, to do the following:

````{r}```
#Adding the tidyr and dplyr libraries
library(tidyr)
library(dplyr)
```` ``` ````

## Task 1

#### Instructions

Load the “iris” dataset and transform it to longer.
- tip: you should get 3 columns Species | Trait | Size

#### Solution

Using `pivot_longer()`.

````{r}```
#Transforming the built-in "iris" dataset to longer
iris_long <- iris %>%
  pivot_longer(
    #Choosing the columns
    cols = c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width),
    #Setting the names of the new columns
    names_to = "Trait",
    values_to = "Size"
   )

#Preview of the first few rows
head(iris_long)
```` ``` ````

## Task 2

#### Instructions 

Answer the following questions.
- tip: you can answer each question “piping” only 2-3 commands

### Task 2.1

What is the mean length and width of the sepal and petal of each species?

#### Solution

Calculating means with summarise grouped by spiecies and traits.

````{r}```
#Transforming the table to show means
 iris_mean <- iris_long  %>% 
  #Grouping by spiecies and traits
  group_by(Species, Trait) %>% 
  #Calculating the means
  summarise(Mean_of_size = mean(Size)) %>% 
  #Sorting
  arrange(Species, Trait)

View(iris_mean)
```` ``` ````
